The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at the European Southern Observatory, Paranal, Chile, Program ID: 088.A-0051 and 091.D-0904.}

### 2017-06-11 GRB111117A.tex
Addresses Lises comments
Closes #8

\abstract{
It is notoriously difficult to localize short $\gamma$-ray bursts (sGRBs) and
their hosts to measure their redshifts. These measurements, however, are
critical to constrain sGRB progenitors and delay time models. Here, we present
spectroscopy of the host galaxy of GRB~111117A and measure its redshift to be $z
= 2.211$. This makes GRB~111117A the most distant high-confidence short GRB
detected to date. Our spectroscopic redshift supersedes a lower redshift value
for this burst previously estimated from photometry.

@@

From the age of the universe at the time of GRB explosion, an initial neutron
star (NS) separation of $a_0 < 3.2~R_\odot$ is required in the case where the
progenitor is a pair of inspiralling NSs. This constraint excludes some of the
longest formation channels for this burst.

@@

There is now mounting evidence that most short-duration $\gamma$-ray bursts
come from the merger of NSs, either with another NS, or a black
hole, due to their apparent association with kilonovae \citep{Barnes2013a,
	Tanvir2013b, Yang2015, Jin2016, Rosswog2016}. The absence of associated
supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b}
supports this idea and distinguishes the physical origin of sGRBs from their
long-duration counterparts. %\citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.

The host galaxies of sGRBs are diverse. They are more massive and less actively 
star-forming on average than lGRB hosts \citep{Fong2013b}, while in some cases, no host 
galaxy can be identified \citep{Berger2010a, Tunnicliffe2014}. Together with their position 
within their hosts \citep{Fong2013a}, this suggests a progenitor system that can be very
long-lived, and is associated with stellar mass rather than star-formation rate (SFR)
\citep{Berger2014}. The median redshift for sGRBs is $z\sim0.5$ \citep{Berger2014}, and
because most of these measurements come from the associated hosts, it is arguably biased
towards lower redshifts.

@@

The total lifetimes of NS binaries depends on their initial separations and
subsequent inspiral times. The delay time from formation to explosion impacts
the timing and distribution of the enrichment of the ISM with heavy $r$-process
elements \citep{VandeVoort2015, Wallner2015,  Ji2016}. Some limits can be
calculated based on models of star-formation histories of, and the spatial
distribution of sGRBs in, their host galaxies \citep[][]{Berger2014}. The most
distant cosmological bursts, however, offer direct, hard limits.

In this \emph{Letter} we present a spectrum of the host galaxy of the short GRB~111117A ($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This value is significantly higher 
than the previously estimated redshift based on photometric studies. 
We present the GRB's rest frame 
properties based on this new distance compared to previous analyses
\citep{Margutti2012,Sakamoto2013} and revisit the host properties derived from
the new solution to the SED fit. %While no optical afterglow was detected, the
% excellent localization from a detection
%of the X-ray afterglow by the \emph{Chandra X-ray Observatory} also allows
%us to discuss the positioning and environmental properties of this remarkably
%distant short GRB.
Throughout this \emph{Letter} we use the $\Lambda$CDM cosmology parameters provided by
\citet{Planck2015} in which the universe is flat with $H_0 = 67.7$\,km\,s$^{-1}$
and $\Omega_m = 0.307$. All magnitudes are given in the AB system.

@@

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs4.pdf}
	\caption{Imaging of the field of GRB~111117A with the X-shooter slit overlaid. Only one slit is shown, despite four epochs of spectroscopic observations because of the similarity in position angle. The image is the FORS2 $R$-band image for which the photometry is shown in Fig.~\ref{fig:SED}. 
	The blue asterisk indicates the GRB position as derived from the \emph{Chandra} observations in \citet{Sakamoto2013}. The insert shows the 2D-image of the \oiii$\lambda$5007 emission line. The location of a bright skyline is marked by the blue box. The location of the emission line is indicated with the red ellipse. Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
	\label{fig:spec_setup}
\end{figure}

Spectroscopic observations were carried out using the cross-dispersed echelle
spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
burst was observed 38 hours after the Burst Alert Telescope (BAT) trigger under
ESO programme 088.A-0051 (PI: Fynbo) and again later under ESO programme
091.D-0904 (PI: Hjorth). X-shooter covers the wavelength range from 3000~\AA~to
24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used) across three
spectroscopic arms. We carried out the bias-correction, flat-fielding, order
tracing, wavelength calibration, rectification, and flux calibration using the
VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010}.
%Because the spectra are curved across each detector, a
% rectification algorithm
%is employed which introduces correlations between neighboring pixels. We select
%a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to minimize the
%degree of correlation while conserving the maximal resolution.
The observations are combined and extracted using scripts described in Selsing
et al. 2017 (in prep.) and available
online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. 
%The
%signal-to-noise of the continuum in the near-infrared arm is too low to use the
%optimal extraction algorithm \citep{Horne1986} and therefore the extraction is
%carried out with a simple aperture.
An overview of the spectroscopic observations is given in
Table~\ref{tab:spec_overview}, and the slit position is shown in
Fig.~\ref{fig:spec_setup}. We show the extracted spectrum in Fig.~\ref{fig:SED}.

We determine a redshift of $z = 2.211$ from the simultaneous detection of
emission lines interpreted as \lya, \oii, \hb, \oiii$\lambda$5007, and \ha. 
\hb~is detected at low significance ($\sim 3 \sigma$). We show \oiii$\lambda$5007
in the insert in Fig.~\ref{fig:spec_setup}.  \ha~is only visible in the first
epoch due to the $K$-band blocking filter used for the remainder observations.
The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and show
significant velocity structure along the slit. A drop in the continuum to the
blue of the \lya~line further supports the inferred redshift.

@@

Using the luminosity of \ha, we can infer the star-formation rate (SFR) of a 
galaxy \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed at 
21\,000~\AA~where the night sky is very bright. In addition, several bright sky-lines 
are superposed on the line, making an accurate estimate of the \ha-flux difficult. 
We obtain a limit on the SFR by integrating the part of \ha~free of contamination and 
correcting for the missing flux using the line shape. After converting the 
\citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial mass function using
\citet{Madau2014}, we derive a limit of $SFR > 7 M_\odot$~yr$^{-1}$. 
From the SED-fit (Sect.~\ref{SED}), and the detection of \lya, the host is
constrained to contain very little or no dust, which is why we do not apply 
a dust-correction to the measured \ha~flux here.

@@

In addition to the spectroscopy presented above, we imaged the field of
GRB~111117A in multiple broad-band filters using the VLT equipped with FORS2
($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters), long after the burst
faded. These new data are complemented by a re-analysis of some of the imaging
used in \citet{Margutti2012} and \citet{Sakamoto2013} that are available to us
(GTC $gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the photometric
observations and measured brightnesses is given in
Table~\ref{tab:phot_overview}.

All data were reduced, analyzed and fitted in a similar manner as described in detail in
\citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}. Briefly, we use our own
\texttt{Python} and IRAF routines to perform a standard reduction which includes
bias/flat-field correction, de-fringing (if necessary), sky-subtraction, and
stacking of individual images. The photometry of the host was tied against magnitudes 
of field stars from the SDSS and 2MASS catalogs in the case of $grizJHK_{\mathrm{s}}$ filters. 
%For the $R$ and $I$-band photometry, we used the color transformations of
%Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. 
We convert all magnitudes into the AB system, and correct for a Galactic 
foreground of $E_{B-V}=0.027~\mathrm{mag}$.

The multi-color spectral energy distribution (SED) is fit by \citet{Bruzual2003}
single stellar population models based on a \citet{Chabrier2003} initial mass
function in \emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to
the spectroscopic value of $z=2.211$. The best fit model is an
unreddened galaxy template, and returns physical parameters of absolute
magnitude ($M_B=-22.0\pm0.1$\,mag), stellar mass ($\log(M_{\star}/M_\odot) =
9.9\pm0.2$), stellar population age ($\tau = 0.5_{-0.3}^{+0.5}$ Gyr) and
star-formation rate ($SFR_{\mathrm{SED}}=11_{-4}^{+9}
M_\odot\,\mathrm{yr}^{-1}$). We show the SED fit in Fig.~\ref{fig:SED}.

Noteworthy is the discrepancy of our new VLT/FORS2 photometry and the
re-analysis of the Gemini data to the $z$-band measurements of \citet{Margutti2012} and
\citet{Sakamoto2013}. Both of these authors report $z$-band photometry that is brighter 
by 0.8~mag to 1.0 mag compared to our value, where data taken in bluer filters are in excellent agreement. The large $i-z$ color is mistakenly interpreted as a 4000\,\AA\,break
driving the galaxy photometric redshift of the earlier works. Using
the revised photometry from Table~\ref{tab:phot_overview}, the photometric redshift of the
galaxy is $z_{\mathrm{phot}}=2.04_{-0.21}^{+0.19}$, consistent with the
spectroscopic value at the 1~$\sigma$ confidence level.

@@

As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A is
securely classified as belonging to the short class of GRBs. Because the
observed classification indicators, $T_{90}$ and hardness ratio, do not depend
strongly on redshift \citep{Littlejohns2013a}, the updated redshift does not
change this designation. %The intrinsic luminosity increases, as
% visible in the X-ray light curve
%in Fig. \ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to
%the majority of long GRBs.
\citet{Bromberg2013} investigated the degree to which the long and short
population distributions overlap and quantified the certainty in the class
membership. According to \citet{Bromberg2013}, GRB~111117A has $96_{-5}^{+3}$
percent probability of being a sGRB. Compared to the other two sGRB candidates
at high redshift, GRB~060121 \citep{DeUgartePostigo2006, Levan2006} at $1.7
\lesssim z \lesssim 4.5$ ($17_{-15}^{+14}$ per cent) and GRB~090426
\citep{Antonelli2009, Levesque2010, Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$
per cent), the certainty in class membership for GRB~111117A is much higher.
This means that GRB~111117A is likely older by 1.8 Gyr than any other
non-collapsar GRB ever detected.

@@

%\begin{figure}
%	\centering
%	\includegraphics[width=9cm]{figures/XLC_111117A_rest.pdf}
%	\caption{Restframe XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs. The grey shared region is a compilation of long GRB lightcurves where the color represents density and the light blue is other short GRB lightcurves for which redshifts has been determined. Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and still subluminous compared to the lGRB population.}
%	\label{fig:sxray_lightcurve}
%\end{figure}

@@

For lGRBs, $N_H$ correlates with the surface luminosity at the explosion site
\citep{Lyman2017}, which is a challenge to reconcile  with the relatively large
offset from the host center derived in \citet{Margutti2012} and
\citet{Sakamoto2013}. This could suggest that there is some galaxy under the
burst position that we don’t see. Along with the absence of dust, the large
offset from the host center indicates that the high $N_H$ arises because the
density in the GRB surrounding is high, or because the light from the afterglow
traverses a region of star formation.

@@

%From the clear host association, GRB~111117A does not belong to the hostless class of GRBs \citep{Berger2010a} and because the host exhibits emission lines indicative of a population of relatively young stars, 
As the majority of sGRBs \citep{Fong2013b}, the host of GRB~111117A is a
late-type galaxy and is entirely consistent in terms of stellar mass and stellar
age with the general population of sGRB hosts ($\left\langle M _*
\right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle =
0.3 $~Gyr) \citep{Leibler2010}. Being a late-type host, both the stellar mass
and sSFR are entirely within the range expected for the hosts of sGRBs
\citep{Behroozi2014}. The SFR is $\sim$1 order of magnitude higher than the
typical SFR for sGRB host galaxies \citep{Berger2014} and more similar to
the SFR found in the hosts of lGRBs at a corresponding redshift
\citep{Kruhler2015}. The high SFR is partly a selection effect, because a less
star forming galaxy would exhibit weaker emission lines, thus making the
redshift harder to determine. Additionally, it is natural to expect some
evolution in the hosts of sGRBs with redshift as illustrated in Sect.~\ref{restnH}.

%The detection of \lya~is consistent with the SED-inferred absence of dust, despite
%the moderate stellar age and the high X-ray derived hydrogen column density
%which would suggest the opposite. The centroid of the Ly$\alpha$ emission is
%found to be redshifted by $\sim$~220~km~s$^{-1}$ with respect to systemic, which
%is similar to what is found for long GRB hosts \citep{Milvang-Jensen2012a} where
%the outflow is attributed to star formation.

@@

If we use the stellar population age from our SED fit, then we obtain a
(softer) limit on the initial separation of $a_0 < 2.1~R_\odot$. However, this
does not account for the possibility there could be an underlying stellar
population of older stars from a previous star-formation epoch. The delay time
between formation and explosion is well accommodated by the models of
\citet{Belczynski2006}, although the longest delay times are excluded.
This is especially true given the late type nature of the host
\citep{OShaughnessy2008}.

@@

Using the new distance, the X-ray derived $n_H$ towards GRB~111117A is the
highest within a complete sample of sGRB hosts and is consistent with the
$n_H-z$ evolution traced by the hosts of lGRBs.The SFR of the host is in the
upper end of the sGRB host SFR distribution and despite the moderate age and
high $N_\mathrm{H}$, almost no dust is present. The high $N_H$ is difficult to
reconcile with the large projected host offset and the absence of dust. One
possible explanation could be, that GRB~111117A is formed through the prompt
channel of sGRBs \todo{ref} and originates in a star forming region located in
the outskirts of the host.

>>>>

\abstract{
It is notoriously difficult to localize short $\gamma$-ray bursts (sGRBs) and
their hosts to measure their redshifts. These measurements, however, are
critical to constrain sGRB progenitors and delay time models. Here, we present
spectroscopy of the host galaxy of GRB~111117A and measure its redshift to be $z
= 2.211$. This makes GRB~111117A the most distant high-confidence short duration GRB
detected to date. Our spectroscopic redshift supersedes a lower redshift value
for this burst previously estimated from photometry.

@@

From the age of the universe at the time of GRB explosion, an initial neutron
star (NS) separation of $a_0 < 3.2~R_\odot$ is required in the case where the
progenitor is a circular pair of inspiralling NSs. This constraint excludes some of the
longest formation channels for this burst.

@@

There is mounting evidence that short-duration $\gamma$-ray bursts
come from the merger of NSs, either with another NS, or a black
hole, due to their apparent association with kilonovae \citep{Barnes2013a,
	Tanvir2013b, Yang2015, Jin2016, Rosswog2016}. The absence of associated
supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b}
supports this idea and distinguishes the physical origin of sGRBs from their
long-duration counterparts. %\citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.

The host galaxies of sGRBs are diverse. They are more massive and less actively
star-forming on average than lGRB hosts \citep{Fong2013b}, while in some cases,
no host galaxy can be identified above detection threshold of deep follow-up
observations \citep{Berger2010a, Tunnicliffe2014}. Together with their position
within their hosts \citep{Fong2013a}, this suggests a progenitor system that can
be very long-lived, and is associated with stellar mass rather than
star-formation rate (SFR) \citep{Berger2014}. The median redshift for sGRBs is
$z\sim0.5$ \citep{Berger2014}, and because most of these measurements come from
the associated hosts, it is arguably biased towards lower redshifts.

@@

The total lifetimes of NS binaries depends on their orbit, mass, spin, initial
separations and subsequent inspiral times. The delay time from formation to
explosion impacts the timing and distribution of the enrichment of the ISM with
heavy $r$-process elements \citep{VandeVoort2015, Wallner2015,  Ji2016}. Some
limits can be calculated based on models of star-formation histories of, and the
spatial distribution of sGRBs in, their host galaxies \citep[][]{Berger2014}.
The most distant cosmological bursts, however, offer direct, hard limits on the
coalescence time scales.

In this \emph{Letter} we present a spectrum of the host galaxy of the short
GRB~111117A ($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This
value is significantly higher than the previously estimated redshift based on
photometric studies \citep{Margutti2012,Sakamoto2013}. We present the GRB's rest
frame properties based on this new distance compared to previous analyses and
revisit the host properties derived from the new solution to the spectral energy distribution (SED) fit.
%While no optical afterglow was detected, the excellent localization from a
%detection of the X-ray afterglow by the \emph{Chandra X-ray Observatory} also
%allows us to discuss the positioning and environmental properties of this
%remarkably distant short GRB.
Throughout this \emph{Letter} we use the $\Lambda$CDM cosmology parameters provided by
\citet{Planck2015} in which the universe is flat with $H_0 = 67.7$\,km\,s$^{-1}$
and $\Omega_m = 0.307$. All magnitudes are given in the AB system.

@@

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs4.pdf}
	\caption{
	Imaging of the field of GRB~111117A with the X-shooter slit overlaid. The slit
	position represents 4 epochs of spectroscopic observations taken at similar
	position angles. The image is the FORS2 $R$-band image for which the photometry
	is shown in Fig.~\ref{fig:SED}. The blue asterisk indicates the GRB position as
	derived from the \emph{Chandra} observations in \citet{Sakamoto2013}. The
	insert shows the 2D-image of the \oiii$\lambda$5007 emission line. The location
	of a bright skyline is marked by the blue box. The location of the emission
	line is indicated with the red ellipse. Because the host is observed in
	nodding-mode, negative images of the emission line appear on both sides in the
	spatial direction.
	}
	\label{fig:spec_setup}
\end{figure}

Spectroscopic observations were carried out using the cross-dispersed echelle
spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
burst was observed 38 hours after the Burst Alert Telescope (BAT) trigger under
ESO programme 088.A-0051 (PI: Fynbo) and again later under ESO programme
091.D-0904 (PI: Hjorth). X-shooter covers the wavelength range from 3000~\AA~to
24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used) across three
spectroscopic arms. We carried out the bias-correction, flat-fielding, order
tracing, wavelength calibration, rectification, and flux calibration using the
VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010}.
%Because the spectra are curved across each detector, a
% rectification algorithm
%is employed which introduces correlations between neighboring pixels. We select
%a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to minimize the
%degree of correlation while conserving the maximal resolution.
The observations are combined and extracted using scripts described in Selsing
et al. 2017 (in prep.) and available
online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. 
%The
%signal-to-noise ratio of the continuum in the near-infrared arm is too low to use the
%optimal extraction algorithm \citep{Horne1986} and therefore the extraction is
%carried out with a simple aperture.
An overview of the spectroscopic observations is given in
Table~\ref{tab:spec_overview}, and the slit position is shown in
Fig.~\ref{fig:spec_setup}. We show the extracted spectrum in Fig.~\ref{fig:SED}.

We determine a redshift of $z = 2.211$ from the simultaneous detection of
emission lines belonging to \lya, \oii, \hb, \oiii$\lambda$5007, and \ha. 
\hb~is detected at low significance ($\sim 3 \sigma$). We show \oiii$\lambda$5007
in the insert in Fig.~\ref{fig:spec_setup}.  \ha~is only visible in the first
epoch due to the $K$-band blocking filter used for the remainder observations.
The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and show
significant velocity structure along the slit. A drop in the continuum to the
blue of the \lya~line further supports the inferred redshift.

@@

Using the luminosity of \ha, we can infer the star-formation rate (SFR) of a 
galaxy \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed at 
21\,000~\AA~where the night sky is very bright. In addition, several bright sky-lines 
are superposed on the line, making an accurate estimate of the \ha-flux difficult. 
We obtain a limit on the SFR by integrating the part of \ha~free of contamination and 
correcting for the missing flux using the line shape. After converting the 
\citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial mass function using
\citet{Madau2014}, we derive a limit of $SFR > 7 M_\odot$~yr$^{-1}$. 
From the SED-fit (Sect.~\ref{SED}), and the detection of \lya, the host is
constrained to contain very little or no dust, although the presence of Lya does
not exclude dust. Therefore we do not apply a dust-correction to the measured
\ha~flux here.

@@

In addition to the spectroscopy presented above, we imaged the field of
GRB~111117A in multiple broad-band filters using the VLT equipped with FORS2
($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters), long after the burst
faded. These new data are complemented by a re-analysis of some of the imaging
used in \citet{Margutti2012} and \citet{Sakamoto2013} that are available to us
(GTC $gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the photometric
observations and measured brightnesses is given in
Table~\ref{tab:phot_overview}. Due to the optical darkness of this burst, there
is likely no afterglow contamination on the measured photometry.

All data were reduced, analyzed and fitted in a similar manner as described in
detail in \citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}.
Briefly, we use our own \texttt{Python} and IRAF routines to perform a standard
reduction which includes bias/flat-field correction, de-fringing (if necessary),
sky-subtraction, and stacking of individual images. The photometry of the host
was calibrated relative to field stars from the SDSS and 2MASS catalogs in the
case of $grizJHK_{\mathrm{s}}$ filters.
%For the $R$ and $I$-band photometry, we used the color transformations of
%Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. 
We convert all magnitudes into the AB system, and correct for a Galactic 
foreground of $E_{B-V}=0.027~\mathrm{mag}$.

The multi-color SED is fit by \citet{Bruzual2003} single stellar population
models based on a \citet{Chabrier2003} with initial mass function in
\emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to the
spectroscopic value of $z=2.211$. The best fit model is an unreddened galaxy
template, and returns physical parameters of absolute magnitude
($M_B=-22.0\pm0.1$\,mag), stellar mass ($\log(M_{\star}/M_\odot) = 9.9\pm0.2$),
stellar population age ($\tau = 0.5_{-0.3}^{+0.5}$ Gyr) and star-formation rate
($SFR_{\mathrm{SED}}=11_{-4}^{+9} M_\odot\,\mathrm{yr}^{-1}$). We show the SED
fit in Fig.~\ref{fig:SED}.

Noteworthy is the discrepancy of our new VLT/FORS2 photometry and the
re-analysis of the Gemini data to the $z$-band measurements of
\citet{Margutti2012} and \citet{Sakamoto2013}. Both of these authors report
$z$-band photometry that is brighter by 0.8~mag to 1.0 mag compared to our
value, where data taken in bluer filters are in excellent agreement. The large
$i-z$ color was mistakenly interpreted as a 4000\,\AA\,break driving the galaxy
photometric redshift of the earlier works. Using the revised photometry from
Table~\ref{tab:phot_overview}, the photometric redshift of the galaxy is
$z_{\mathrm{phot}}=2.04_{-0.21}^{+0.19}$, consistent with the spectroscopic
value at the 1~$\sigma$ confidence level.

@@

As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A is
securely classified as belonging to the short class of GRBs. Because the
observed classification indicators, $T_{90}$ and hardness ratio, do not depend
strongly on redshift \citep{Littlejohns2013a}, the updated redshift does not
change this designation.
%The intrinsic luminosity is higher as demonstrated in the X-ray light curve
%(Fig. \ref{fig:sxray_lightcurve}), but it is still is sub-luminous compared to
%the majority of long GRBs.
\citet{Bromberg2013} investigated the degree to which the long and short
population distributions overlap and quantified the certainty in the class
membership. According to \citet{Bromberg2013}, GRB~111117A has $96_{-5}^{+3}$
percent probability of being a sGRB. Compared to the other two sGRB candidates
at high redshift, GRB~060121 \citep{DeUgartePostigo2006, Levan2006} at $1.7
\lesssim z \lesssim 4.5$ ($17_{-15}^{+14}$ per cent) and GRB~090426
\citep{Antonelli2009, Levesque2010, Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$
per cent), the certainty in class membership for GRB~111117A is much higher.
This means that GRB~111117A happened when the universe was younger by 1.8 Gyr
compared to any other non-collapsar GRB ever detected. This number is 3.2 Gyr
for the next-highest spectroscopic redshift. This has implications for the
$r$-process element enrichment timescale of the universe.

@@

%\begin{figure}
%	\centering
%	\includegraphics[width=9cm]{figures/XLC_111117A_rest.pdf}
%	\caption{Restframe XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs. The grey shaded region is a compilation of long GRB lightcurves from \todo{reference} where the color represents density and the light blue lines are other short GRB lightcurves from \todo{reference} for which redshifts has been determined. Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and still subluminous compared to the lGRB population.}
%	\label{fig:sxray_lightcurve}
%\end{figure}

@@

%For lGRBs, $N_H$ correlates with the surface luminosity at the explosion site
%\citep{Lyman2017}, which is a challenge to reconcile  with the relatively large
%offset from the host center derived in \citet{Margutti2012} and
%\citet{Sakamoto2013}. This could suggest that there is some galaxy under the
%burst position that we don’t see. Along with the absence of dust, the large
%offset from the host center indicates that the high $N_H$ arises because the
%density in the GRB surrounding is high, or because the light from the afterglow
%traverses a region of star formation.

@@

%From the clear host association, GRB~111117A does not belong to the hostless
%class of GRBs \citep{Berger2010a} and because the host exhibits emission lines
%this is indicative of a population of relatively young stars.
As the majority of sGRBs \citep{Fong2013b}, the host of GRB~111117A is a
late-type galaxy and is entirely consistent in terms of stellar mass and stellar
age with the general population of sGRB hosts ($\left\langle M _*
\right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle =
0.3 $~Gyr) \citep{Leibler2010}. Being a late-type host, both the stellar mass
and sSFR are entirely within the range expected for the hosts of sGRBs
\citep{Behroozi2014}. The SFR is $\sim$1 order of magnitude higher than the
typical SFR for sGRB host galaxies \citep{Berger2014} and more similar to
the SFR found in the hosts of lGRBs at a corresponding redshift
\citep{Kruhler2015}. The high SFR is partly a selection effect, because a less
star forming galaxy would exhibit weaker emission lines, thus making the
redshift harder to determine. Additionally, it is natural to expect some
evolution in the hosts of sGRBs with redshift as illustrated in Sect.~\ref{restnH}.

%The detection of \lya~is consistent with the SED-inferred absence of dust,
%despite the moderate stellar age which would suggest the opposite. The centroid
%of the Ly$\alpha$ emission is found to be redshifted by $\sim$~220~km~s$^{-1}$
%with respect to systemic, which is similar to what is found for long GRB hosts
%\citep{Milvang-Jensen2012a} where the outflow is attributed to star formation.

@@

Using the inferred stellar population age from our SED fit, then we obtain a
(softer) limit on the initial separation of $a_0 < 2.1~R_\odot$. However, this
does not account for the possibility there could be an underlying stellar
population of older stars from a previous star-formation epoch. The delay time
between formation and explosion is well accommodated by the models of
\citet{Belczynski2006}, although the longest delay times are excluded.
This is especially true given the late type nature of the host
\citep{OShaughnessy2008}.

@@

Using the new distance, the X-ray derived $n_H$ towards GRB~111117A is the
highest within a complete sample of sGRB hosts and is consistent with the
$n_H-z$ evolution traced by the hosts of lGRBs.The SFR of the host is in the
upper end of the sGRB host SFR distribution and no dust is present. The high
$N_H$ is difficult to reconcile with the large projected host offset and the
absence of dust. One possible explanation could be, that GRB~111117A is formed
through the prompt channel of sGRBs \todo{ref} and originates in a star forming
region located in the outskirts of the host.
